Fixed-Parameter Tractable Sampling for RNA Design with Multiple Target Structures

### 2018-08-31 supplement.tex
Write clean proof of hyper-graph tree decomposition

\usepackage{xr}
\externaldocument{RNARedPrint-Bioinfo2018}

@@

%\newtheorem{algorithm}{Algorithm}
\newtheorem{axiom}{Axiom}
\newtheorem{hypothesis}{Working Hypothesis}
\newenvironment{proof}[1][]{\noindent \em Proof\ifthenelse{\equal{#1}{}}{}{ (#1)}:~}{}

@@

Crucially, one can show that, as long as the arguments $\dep(f)$ of a function $f$ form a clique within the network $\network$, then there exists at least one node in the tree decomposition that contains $\dep(f)$, thus allowing the evaluation of $f$.
\begin{lemma}\label{lem:cliques}
\todo[inline]{R1: guess it should read "node v" instead of "n"!? please update the proof accordingly!}
Let $G=(V,E)$ be an undirected graph, and  $(T,\chi)$ be a tree decomposition for $G$. For each clique $C\subset V$, there exists a node $n\in T$ such that $C\subset\chi(v)$.
\end{lemma}
\begin{proof}
  \newcommand{\TransVars}[1]{{\rm tverts}(#1)}
  \newcommand{\Children}[1]{{\rm Children}(#1)}
  For the sake of simplicity, let us consider $T$ as rooted on an arbitrary node, inducing an orientation, and denote by $\TransVars{n}\in V$ the set of vertices found in a node $n$ or its descendants, namely
    $$\TransVars{n} = \chi(n) \bigcup_{n'\in\Children{n}} \TransVars{n'}.$$

  Now consider (one of) the deepest node(s) $n\in T$, such that $C\subset\TransVars{v}$. We are going to prove that $C\subset\chi(n)$, using contradiction, by showing that $C\not\subset\chi(n)$ implies that $T$ is not a tree decomposition.

  Indeed, let us assume that $C\not\subset\chi(n)$, then there exists a node $v'\notin\chi(n)$ whose presence in $\TransVars{n}$ stems from its presence in some descendant of $n$. Let us denote by $n'\neq n$ the closest descendant of $n$ such that $v'\in\chi(n')$. Now, consider a node $v\in C$ such that $v\in \TransVars{v'}$ and $v\notin \TransVars{v'}$. Such a node always exists, otherwise $n'$, and not $n$, would be the deepest node such that $C\subset \TransVars{v'}$. From the definition of the tree decomposition, we know that neither $n'$ nor its descendants may contain $v$. On the other hand, none of the parents or siblings of $n'$ may contain $v'$. It follows that is no node $n''\in T$ whose vertex set $\chi(v'')$ includes, at the same time, $v$ and $v'$. Since both $v$ and $v'$ belong to a clique, one has $\{v,v'\}\in E$. The absence of a node in $T$ capturing an edge in $E$ contradicts our initial assumption that $T$ is tree decomposition for $G$.

  We conclude that $n$ is such that $C\not\subset\chi(n)$.
\end{proof}

@@

\bibliographystyle{plain}
%\bibliographystyle{achemnat}
%\bibliographystyle{plainnat}
%\bibliographystyle{abbrv}
%\bibliographystyle{bioinformatics}
%
%\bibliographystyle{plain}
%
\bibliography{biblio}

>>>>

\usepackage{xr}
\externaldocument{RNARedPrint}

@@

%\newtheorem{algorithm}{Algorithm}
\newtheorem{axiom}{Axiom}
\newtheorem{hypothesis}{Working Hypothesis}
\newenvironment{proof}[1][]{\noindent \emph{Proof}\ifthenelse{\equal{#1}{}}{}{ (#1)}.~}{\hfill$\Box$}

@@

Crucially, one can show that, as long as the arguments $\dep(f)$ of a function $f$ form a clique within the network $\network$, then there exists at least one node in the tree decomposition that contains $\dep(f)$, thus allowing the evaluation of $f$.
%
While this result is certainly part of folklore, we emphasize the importance for the generality of this work and present a brief direct proof for convenience.

\begin{lemma}\label{lem:cliques}
Let $G=(V,E)$ be an undirected graph, and  $(T,\chi)$ be a tree decomposition for $G$. For each clique $C\subseteq V$, there exists a node $u \in T$ such that $C\subseteq\chi(u)$.
\end{lemma}

%\begin{proof}
% \newcommand{\tn}{v_{\rm T}} %% node in tree
% \newcommand{\TransVars}[1]{{\rm tverts}(#1)}
% \newcommand{\Children}[1]{{\rm Children}(#1)}
%  For the sake of simplicity, let us consider $T$ as rooted on an arbitrary node, inducing an orientation, and denote by $\TransVars{\tn}\in V$ the set of vertices found in the node $\tn$ or its descendants, namely
%    $$\TransVars{\tn} = \chi(\tn) \cup \bigcup_{\tn'\in\Children{\tn}} \TransVars{\tn'}.$$
%
%  Now consider (one of) the deepest node(s) $\tn\in T$, such that $C\subseteq\TransVars{\tn}$. We are going to prove that $C\subseteq\chi(\tn)$, using contradiction, by showing that $C\not\subseteq\chi(\tn)$ implies that $T$ is not a tree decomposition.
%
%  Assume that $C\not\subseteq\chi(\tn)$, then there exists a node $v'\notin\chi(\tn)$ whose presence in $\TransVars{\tn}$ stems from its presence in some descendant of $\tn$. Let us denote by $\tn'\neq \tn$ the closest descendant of $\tn$ such that $v'\in\chi(\tn')$. Now, consider a node $v\in C$ such that $v\in \TransVars{\tn'}$ and $v\notin \TransVars{\tn'}$. Such a node always exists, otherwise $\tn'$, and not $\tn$, would be the deepest node such that $C\subseteq \TransVars{\tn'}$. From the definition of the tree decomposition, we know that neither $\tn'$ nor its descendants may contain $v$. On the other hand, none of the parents or siblings of $\tn'$ may contain $v'$. It follows that is no node $\tn''\in T$ whose vertex set $\chi(\tn'')$ includes, at the same time, $v$ and $v'$. Since both $v$ and $v'$ belong to a clique, one has $\{v,v'\}\in E$. The absence of a node in $T$ capturing an edge in $E$ contradicts our initial assumption that $T$ is tree decomposition for $G$.
%
%  We conclude that $\tn$ is such that $C\not\subseteq\chi(\tn)$.
%\end{proof}

\begin{proof}
  \newcommand{\TransVars}[1]{{\rm tverts}(#1)}%
  Root $T$ at an arbitrary tree node $r\in T$, which induces a parent/child relation between tree nodes.
  We say a tree node $u$ \emph{represents} the node $v\in V$ (respectively, the set of nodes $X\subseteq V$) iff $v\in\chi(u)$ ($X\subseteq\chi(u)$).
  Moreover, denote by $\TransVars{u}\subseteq V$ the vertices
  represented by the nodes of the subtree of $u$, i.e.
  $$\TransVars{u} := \chi(u) \cup \bigcup_{\text{children $u'$ of u}} \TransVars{u'}.$$

  Let $u$ denote the node in $T$ with maximal distance to the root $r$,
  such that still $C\subseteq\TransVars{u}$.
  Let $C'$ be $C-\chi(u)$.
  If $C'$ is empty than $u$ represents $C$---nothing remains to be shown;
  thus assume, $C' \neq \emptyset$.
  For exactly one of the children $u'$ of $u$,
  $C'\subseteq\TransVars{u'}$, since otherwise for
  some node $v'\in C'$, the subtree of the nodes representing $v'$ would be disconnected (property 2 of Def.~\ref{def:treedecomp}). Now, pick some $v'\in C'$ and some
  $v''\in C-C'$, which must be represented by $u$ or some child $u''$ of $u$. In both cases, there cannot
  be any node in T that represents the edge $\{v',v''\}\subseteq C$, since $v'$ can only be represented 
  by nodes in the subtree of $u'$.
\end{proof}

@@

\bibliographystyle{vancouver}
%\bibliographystyle{achemnat}
%\bibliographystyle{plainnat}
%\bibliographystyle{abbrv}
%\bibliographystyle{bioinformatics}
%
%\bibliographystyle{plain}
%
\bibliography{biblio}
